---
title: "DATA MEMO"
author: "Zhimei_Chen"
date: '2022-12-11'
output: 
  html_document:
    toc: true
    toc_depth: 4
    toc_float: true
    code_folding: show
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Libraries Needed: 
```{r, warning=FALSE,message=FALSE,echo=FALSE}
library(dplyr)
library(dbplyr)
library(tidyverse) 
library(RColorBrewer)
library(ggplot2)
library(corrplot)
library(ISLR)
library(tidymodels)
library(glmnet)
library(rpart)
library(rpart.plot)
library(randomForest)
library(ggpubr)
library(kernlab)
library(ranger)
library(class)
library(FNN)
library(vip)
tidymodels_prefer()
```

### DATA 
#### Credit Card Approval Prediction\
https://www.kaggle.com/datasets/rikdifos/credit-card-approval-prediction?select=application_record.csv \
application_record.csv contains appliers personal information, which I use as features for predicting.\
credit_record.csv records users' behaviors of credit card. \

### read csv and clean
```{r,warning=FALSE,echo=FALSE}
options(readr.show_col_types = FALSE)
options(tibble.width = Inf) 
library(janitor)
application <- read_csv("application_record.csv")%>% 
  clean_names()
credit <- read_csv("credit_record.csv")%>% 
  clean_names()
head(application)
```

### DATA: application_data

#### Observations
```{r}
cat("Observation number in application: ", nrow(application))
```

#### turn to characters and factors and see if there are NAs
```{r,cols.print=18,echo=FALSE}
application2 <- application %>% 
  mutate(id = as.character(id),
         code_gender = factor(code_gender),
         flag_own_car = factor(flag_own_car),
         flag_own_realty = factor(flag_own_realty),
         name_income_type = factor(name_income_type),
         name_education_type = factor(name_education_type),
         name_family_status = factor(name_family_status),
         name_housing_type = factor(name_housing_type),
         flag_mobil = factor(flag_mobil),
         flag_work_phone = factor(flag_work_phone),
         flag_phone = factor(flag_phone),
         flag_email = factor(flag_email),
         occupation_type = factor(occupation_type))

summary(application2)
```
The only variable with missingness is occupation_type, which is missing 134203 observations.

#### Trans NA to "Dont wanna tell"
```{r}
application$occupation_type[is.na(application$occupation_type)] <- "Don't wanna tell"
```

#### turn to characters and factors 
```{r,cols.print=18,echo=FALSE}
application <- application %>% 
  mutate(id = as.character(id),
         code_gender = factor(code_gender),
         flag_own_car = factor(flag_own_car),
         flag_own_realty = factor(flag_own_realty),
         name_income_type = factor(name_income_type),
         name_education_type = factor(name_education_type),
         name_family_status = factor(name_family_status),
         name_housing_type = factor(name_housing_type),
         flag_mobil = factor(flag_mobil),
         flag_work_phone = factor(flag_work_phone),
         flag_phone = factor(flag_phone),
         flag_email = factor(flag_email),
         occupation_type = factor(occupation_type))
head(application)
```


#### Variables
**id**:	Client number\
**code_gender**: Gender [M,F]\
**flag_own_car**: Is there a car [N,Y]\
**flag_own_realty**: Is there a property [N,Y]\
**cnt_children**: Number of children\
**amt_income_total**: Annual income\
**name_income_type**: Income category [Commercial associate, Pensioner, State servant,  Student, Working]\
**name_education_type**: Education level [Academic degree, Higher education, Incomplete higher, Lower secondary, Secondary / secondary special]\
**name_family_status**: Marital status [Civil marriage, Married, Separated, Single / not married, Widow]\
**name_housing_type**: Way of living [Co-op apartment, House / apartment, Municipal apartment, Office apartment, Rented apartment, With parents]\
**days_birth**: Birthday	Count backwards from current day (0), -1 means yesterday\
**days_employed**: Start date of employment	Count backwards from current day(0). If positive, it means the person currently unemployed.\
**flag_mobile**: Is there a mobile phone [0,1]	\
**flag_work_phone**: Is there a work phone [0,1]\
**flag_phone**: Is there a phone	[0,1]\
**flag_email**: Is there an email	[0,1]\
**occupation_type**: Occupation	[Laborers, Core staff, Sales staff, Managers, Drivers, (Other) , NA's]\
**cnt_fam_members**: Family size	\

#### trans inappropriate days of employees to 0 days
```{r}
application$days_employed[application$days_employed==0] ##no one get 0 days
head(application$days_employed[application$days_employed>0])
#Therefore, I think who don't have work has assigned as "365243" days
#trans the number of inappropriate days of employees to 0 days
application$days_employed[application$days_employed == 365243] <- 0
```

### DATA: credit_data
```{r}
head(credit)
summary(credit)
```
There is no missing data in credit.

#### Variables
id: Client number	\
months_balance: Record month	The month of the extracted data is the starting point, backwards, 0 is the current month, -1 is the previous month, and so on\
status: Status (*more in "transfer status"*)


#### Observations
```{r}
cat("Observation number in credit: ",nrow(credit))
```
#### transfer status 
0: 1-29 days past due --> 1\
1: 30-59 days past due --> 2\
2: 60-89 days overdue --> 3\
3: 90-119 days overdue --> 4\
4: 120-149 days overdue --> 5\
5: Overdue or bad debts, write-offs for more than 150 days --> 6\
C: paid off that month --> 0\
X: No loan for the month --> 0
```{r,echo=FALSE}
credit_needed <- credit
credit_needed$id <- as.character(credit_needed$id)
credit_needed$status<- as.character(credit_needed$status)

credit_needed$status[credit_needed$status == "5"] <- 6
credit_needed$status[credit_needed$status == "4"] <- 5
credit_needed$status[credit_needed$status == "3"] <- 4
credit_needed$status[credit_needed$status == "2"] <- 3
credit_needed$status[credit_needed$status == "1"] <- 2
credit_needed$status[credit_needed$status == "0"] <- 1
credit_needed$status[credit_needed$status == "C"] <- 0
credit_needed$status[credit_needed$status == "X"] <- 0

head(credit_needed)
```

#### turn to factors
```{r,cols.print=18}
credit_needed <- credit_needed %>% 
  mutate(status = factor(status))
head(credit_needed)
```

first year credit
```{r}
first_year_credit <- credit_needed %>%
             arrange(id,months_balance,status) %>%
             group_by(id) %>%
             slice(1:12) %>%
             mutate(month = seq(1:length(id)))
           
head(first_year_credit)
```

### Combine two datasets

#### How we combine?
Example

An example here who, with ID: 5008805, applied credit card and had been approved.
```{r,warning=FALSE}
first_year_credit[first_year_credit$id == 5008805,]
credit[credit$id == 5008805,]
application[application$id == 5008805,]
```

Therefore, we can combine application data with first year, first month credit data
```{r}
length(unique(first_year_credit$id)) 
length(unique(credit$id))
length(unique(application$id))

has_credit <- merge(application, first_year_credit, 
                  by.x = "id", by.y = "id")
cat(length(unique(has_credit$id)),"people has credit card with Info here.")

first_month <- has_credit %>%
  filter(month == min(month))
```

#### Observe - graphs

those had been approved's information (numbers)
```{r,echo=FALSE}
numchr <- first_month %>%
  ggplot(aes(x = cnt_children)) +
  geom_boxplot(fill="darkturquoise")+
  ggtitle("Amount of children")
income <- first_month %>%
  ggplot(aes(x = amt_income_total)) +
  geom_boxplot(fill="darkturquoise")+
  ggtitle("Annual income")
age <- first_month %>%
  ggplot(aes(x = abs(days_birth)/365))+
  geom_boxplot(fill="darkturquoise")+
  ggtitle("Age")
employedyear <- first_month %>%
  ggplot(aes(x = abs(days_employed)/365))+
  geom_boxplot(fill="darkturquoise")+
  ggtitle("Years been Employeed")
ctfmem <- first_month %>%
  ggplot(aes(x = cnt_fam_members)) +
  geom_boxplot(fill="darkturquoise")+
  ggtitle("Amount of family members")
ggarrange(numchr, income, age, employedyear, ctfmem, 
          nrow = 3,ncol=2,
          labels = c('a', 'b','c','d','e'))
```

those had been approved's information (factors - YES/NO)
```{r,echo=FALSE}
gender <- first_month %>%
  ggplot(aes(x = code_gender)) +
  geom_bar(fill="darkturquoise", stat = "count") +
  ggtitle("Gender")
iscar <- first_month %>%
  ggplot(aes(x = flag_own_car)) +
  geom_bar(fill="darkturquoise", stat = "count")+
  ggtitle("Has car?")
isrealty <- first_month %>%
  ggplot(aes(x = flag_own_realty)) +
  geom_bar(fill="darkturquoise", stat = "count")+
  ggtitle("Has realty?")
ismobile <- first_month %>%
  ggplot(aes(x = flag_mobil)) +
  geom_bar(fill="darkturquoise", stat = "count")+
  ggtitle("Has mobile?")
isworkphone <- first_month %>%
  ggplot(aes(x = flag_work_phone)) +
  geom_bar(fill="darkturquoise", stat = "count")+
  ggtitle("Has work phone?")
isphone <- first_month %>%
  ggplot(aes(x = flag_phone)) +
  geom_bar(fill="darkturquoise", stat = "count")+
  ggtitle("Has phone?")
isemail <- first_month %>%
  ggplot(aes(x = flag_email)) +
  geom_bar(fill="darkturquoise", stat = "count")+
  ggtitle("Has email?")

ggarrange(gender, iscar, isrealty, ismobile,isworkphone, isphone, isemail,
          nrow = 2,ncol=4,
          labels = c('a', 'b','c','d','e','f','g'))
```

those had been approved's information (factors - REST)
```{r,echo=FALSE}
inctype <- first_month %>%
  ggplot(aes(x = name_income_type)) +
  geom_bar(fill="darkturquoise", stat = "count") +
  coord_flip() +
  ggtitle("Name Income Type")
inctype
edutype <- first_month %>%
  ggplot(aes(x = name_education_type)) +
  geom_bar(fill="darkturquoise", stat = "count") +
  coord_flip() +
  ggtitle("Name Education Type")
edutype
houtype <- first_month %>%
  ggplot(aes(x = name_housing_type)) +
  geom_bar(fill="darkturquoise", stat = "count") +
  coord_flip() +
  ggtitle("Name Housing Type")
houtype
occtype <- first_month %>%
  ggplot(aes(x = occupation_type)) +
  geom_bar(fill="darkturquoise")+
  coord_flip() +
  ggtitle("Occupation Type")
occtype
```

#### Compare numbers of ID
```{r}
cat(length(unique(first_year_credit$id)), 
length(unique(credit$id)),
length(unique(application$id)),
length(unique(first_month$id)))
```
##### My understanding to this data is according to the 45985 people who have credit cards, 36457 people from those 45985 people are applied with recording the backgrounds. The bank issued credit cards to those 36457 people from 438510 applicants.

#### Combine and add our y
```{r,warning=FALSE}
mydata1<-application%>%
  filter(! id %in% first_month$id)%>%
  mutate(get_credit_card=FALSE)
mydata2<-application%>%
  filter(id %in% first_month$id)%>%
  mutate(get_credit_card=TRUE)
mydata<-rbind(mydata1,mydata2)
summary(mydata$get_credit_card)
```
There are 402053 people who do not have credit card, while the other 36457 people have.

```{r,warning=FALSE}
cat(length(unique(mydata$id)),"is less than",nrow(mydata))
```
Finding out there are duplicated rows in mydata, since when unique id should be 43510, where we have 438557 rows of mydata, so I remove them: 

#### Remove duplicated rows & turn logistics to factors
```{r}
mydata<-mydata[!duplicated(mydata$id),]
mydata$get_credit_card<-as.factor(mydata$get_credit_card)
```

### An overview of your dataset
What does it include?\
There are two datasets, one is about the information of the applicants who try to get credit card, and the other one is about people's credit card information. Variables are showing above.\
Where and how will you be obtaining it? Include the link and source.\
About how many observations? How many predictors?\
I will combine two datasets together. My understanding to this data is according to the 45985 people who have credit cards, 36457 people from those 45985 people are applied with recording the backgrounds. The bank issued credit cards to those 36457 people from 438510 applicants. 438510 observations and 18 predictors.\
What types of variables will you be working with?\
Numeric, strings (charactors) and factors.\
Is there any missing data? About how much? Do you have an idea for how to handle it?\
Yes, there are some missing data. The only variable with missingness is occupation_type, which is missing 134203 observations. As it was a string type, so I transfer it to "Dont wanna tell". There are also many people has 0 employee_day but with number 365243. \

### An overview of your research question(s)
What variable(s) are you interested in predicting? What question(s) are you interested in answering? \
The people in credit dataset are who got credit card, where people in application are who applied credit cards. I want to predict what kind of person can successfully get the credit card after application. I want to know which predictor work best.\
Name your response/outcome variable(s) and briefly describe it/them.
Will these questions be best answered with a classification or regression approach?\
My response/outcome variable(s) is get_credit_card, it is binary. TRUE means that person successfully applied the credit card, and FALSE means the person fail to be approved. I think these questions be best answered with a classification or regression approach.\
Which predictors do you think will be especially useful?\
I think code_gender,flag_own_car,flag_own_realty,amt_income_total,name_income_type, name_education_type,name_family_status,name_housing_type,cnt_fam_members are important, where the amount of income are most important.\
Is the goal of your model descriptive, predictive, inferential, or a combination? Explain.
It is combination. It is predictive as I want to make models predict the TRUE or False for me. And I will also use random forest model to get the best predictor.

### Your proposed project timeline
When do you plan on having your data set loaded, beginning your exploratory data analysis, etc? Provide a general timeline for the rest of the quarter.\
Before November 15th to finish KNN model and logistic Regression.
Before the end of November to finish one or more model (include random forest).
Finish the rest models and write introduction and conclusion before final week.


### Any questions or concerns
Are there any problems or difficult aspects of the project you anticipate?\
The datasets are too big. It will take time to run for some models.
Any specific questions you have for me/the instructional team?\
I do not have any questions yet.
